---
output: md_document
---

# Summary of treatment and genotype differences from the retention session on the last day

```{r setup, echo = F, message=F}
library(dplyr)
library(ggplot2)
library(tidyr)
library(cowplot)
library(car)

knitr::opts_chunk$set(fig.path = '../figures/01_retention/')

colorvalAPA2 <-  c( "#404040","#ca0020", "#bababa", "#f4a582")
# dark grey, red, light grey, peach

behavior <- read.csv("../results/behaviordata.csv")
behavior$APA2 <- factor(behavior$APA2, levels = c("yoked-consistent","consistent", "yoked-conflict","conflict")) ## relevel then rename factors treatment
behavior$Genotype <- factor(behavior$Genotype, levels = c("WT","FMR1KO")) # relevel genotype

retention <- filter(behavior, TrainSession == "Retention") 
retention <- retention[c(1,2,8,4,13:51)]
write.csv(retention, "../results/01_retention.csv", row.names = F)
```


## Groups ananlyzed individually

```{r, echo = F}
# This is my function to plot specific variables for spefic groups

groupspecific <- function(group, colorvals, cols){
df1 <- filter(behavior, TrainSession == "Retention", APA2 == group) %>% droplevels()
#table(df1$APA2, df1$Genotype)

for(i in names(df1)[cols]) {
  df2 <- df1[, c(i, "APA2", "Genotype")]
  df2$wrap <- i
  
  plots <- ggplot(df2) + geom_boxplot(aes_string(y = i, x = "Genotype", fill="APA2", color="Genotype")) +   
    theme_cowplot(font_size = 10, line_size = 0.25) +
    scale_fill_manual(values = colorvals) +
   scale_color_manual(values = c("black", "grey")) +
    facet_wrap(~wrap) + background_grid(major = "y", minor = "none")
  
  #png(paste0("../figures/01_retention/", i, ".png", sep = ""), width = 300, height = 170)
  #print(plots)
  #dev.off()
  
  print(i)
  
  colnames(df2)[1] <- "variable"
 
  print(summary(aov(variable ~ Genotype, data= df2)))
  print(t.test(variable ~ Genotype, data= df2))
  print(plots)
}
}
```

### Yoked consistent

**RayleigAngle** is significantly different between WT and FMR1-KO when examining the yoked-consistent group in isolation (ANOVA and t-test). We have RNA sequencing data to compare with this behavioral data. 

```{r, echo=F}
groupspecific("yoked-consistent", c("#404040","#404040"), c(31))
```

### Consistent
**Speed2ndEntr and Max50.RngLoBin** are significantly different between WT and FMR1-KO when examining the consistent group in isolation (ANOVA and t-test).

```{r, echo=F}
groupspecific("consistent", c("#ca0020","#ca0020"), c(24,41))
```

### Yoked conflict

**Speed2ndEntr amd RayleigAngle** are significantly different between WT and FMR1-KO when examining the yoked conflict group in isolation (ANOVA and t-test).

```{r, echo=F}
groupspecific("yoked-conflict", c("#bababa","#bababa"), c(24,31))
```

### Conflict

**Speed2ndEntr** is significantly different between WT and FMR1-KO when examining the conflict group in isolation (ANOVA and t-test).

```{r, echo=F}
groupspecific("conflict", c("#f4a582","#f4a582"), c(24))
```

### Multiple  groups analyzed at once

```{r, echo = F}
# This is my function for looking at multiple groups at a time
multiplegroups <- function(group, colorvals, cols){

df1 <- filter(behavior, TrainSession == "Retention", APA2 %in% group) %>% droplevels()
#print(table(df1$APA2, df1$Genotype))

for(i in names(df1)[cols]) {
  df2 <- df1[, c(i, "APA2", "Genotype")]
  df2$wrap <- i
  
  plots <- ggplot(df2) + geom_boxplot(aes_string(y = i, x = "APA2", fill="APA2", color="Genotype")) +   
    theme_cowplot(font_size = 10, line_size = 0.25) +
    scale_fill_manual(values = colorvals) +
   scale_color_manual(values = c("black", "grey")) +
    facet_wrap(~wrap) + background_grid(major = "y", minor = "none")

    print(i)
  
  colnames(df2)[1] <- "variable"
 
  print(summary(aov(variable ~ Genotype *APA2, data= df2)))
  #print(TukeyHSD(aov(variable ~ Genotype *APA2, data= df2)))
  print(plots)
}
}
```

When controlling for training, Speed2ndEntr, RayleigAngle, and Min50.RngHiBin are significantly different between WT and FMR1-KO.  Number of entrances and path to 1st entrance are not.

```{r, echo = F}
multiplegroups(c("conflict", "consistent","yoked-conflict", "yoked-consistent" ), colorvalAPA2, c(24,31,37,15:17))
``` 

```{r, echo = F}
# Add this to the for loop for prining a png file
  #png(paste0("../figures/01_retention/", i, ".png", sep = ""), width = 300, height = 170)
  #print(plots)
  #dev.off()
```

