---
title: "DESeq2 Data Analysis"
author: "Rayna M Harris"
date: "January 23, 2017"
output:
  md_document:
    variant: markdown_github
---

## RNAseq gene expression analysis with DESeq2 

This workflow was modified from the DESeq2 tutorial found at: https://www.bioconductor.org/packages/release/bioc/vignettes/DESeq2/inst/doc/DESeq2.pdf

First I load a handful of packages for data wrangling, gene expression analysis, data visualization, and statistics.
```{r setup, message=F, warning=F}
library(dplyr) ## for filtering and selecting rows
library(plyr) ## for renmaing factors
library(reshape2) ## for melting dataframe
library(DESeq2) ## for gene expression analysis
library(edgeR)  ## for basic read counts status
library(magrittr) ## to use the weird pipe
library(genefilter)  ## for PCA fuction
library(ggplot2)
library(cowplot) ## for some easy to use themes
library(car) ## stats
library(VennDiagram) ## venn diagrams
library(pheatmap) ## awesome heatmaps
library(viridis) # for awesome color pallette
library(magrittr) ## to use the weird pipe
library(ggrepel) ## for labeling volcano plot

## load functions 
source("figureoptions.R")
source("functions_RNAseq.R")

## set output file for figures 
knitr::opts_chunk$set(fig.path = '../figures/02_RNAseq/')

```

We are ready to calculate differential gene expression using the DESeq package. For simplicity, I will use the standard nameing of "countData" and "colData" for the gene counts and gene information, respectively.

```{r prepforDESeq2}
colData <- read.csv("../data/fmr1ColData.csv", header = T)
countData <- read.csv("../data/fmr1CountData.csv", header = T, check.names = F, row.names = 1)

## remove outliers
colData <- colData %>% 
  filter(RNAseqID != "16-123B")  %>% 
  filter(RNAseqID != "16-125B") %>% 
  droplevels()

savecols <- as.character(colData$RNAseqID) 
savecols <- as.vector(savecols) 
countData <- countData %>% dplyr::select(one_of(savecols)) 

# colData must be factors
cols = c(1:6)
colData[,cols] %<>% lapply(function(x) as.factor(as.character(x)))

# summary data
colData %>% select(Genotype, APA)  %>%  summary()
```

## Total Gene Counts Per Sample 
this could say something about data before normalization

```{r totalRNAseqcounts}
## stats
dim(countData)
counts <- countData
dim( counts )
colSums( counts ) / 1e06  # in millions of reads
table( rowSums( counts ) )[ 1:30 ] # Number of genes with low counts

rowsum <- as.data.frame(colSums( counts ) / 1e06 )
names(rowsum)[1] <- "millioncounts"
rowsum$sample <- row.names(rowsum)

ggplot(rowsum, aes(x=millioncounts)) + 
  geom_histogram(binwidth = 1, colour = "black", fill = "darkgrey") +
  theme_classic() +
  scale_x_continuous(name = "Millions of Gene Counts per Sample",
                     breaks = seq(0, 8, 1),
                     limits=c(0, 8)) +
  scale_y_continuous(name = "Number of Samples")
```

## DeSeq2

```{r DESeq2, echo = T}
## create DESeq object using the factor Genotyp
dds <- DESeqDataSetFromMatrix(countData = countData,
                              colData = colData,
                              design = ~ Genotype)

dds$Genotype <- factor(dds$Genotype, levels=c("WT", "FMR1")) ## specify the factor levels 

dds # view the DESeq object - note numnber of genes
dds <- dds[ rowSums(counts(dds)) > 1, ]  # Pre-filtering genes with 0 counts
dds # view number of genes afternormalization and the number of samples
dds <- DESeq(dds) # Differential expression analysis
rld <- rlog(dds, blind=FALSE) ## log transformed data
rlddf <- assay(rld)

vsd <- getVarianceStabilizedData(dds)
```

## PCA

```{r pca, echo = T, message=F, warning=F}
# create the dataframe using my function pcadataframe
pcadata <- pcadataframe(rld, intgroup=c("Genotype"), returnData=TRUE)
percentVar <- round(100 * attr(pcadata, "percentVar"))
percentVar

aov1 <- aov(PC1 ~ Genotype, data=pcadata)
summary(aov1) 

aov2 <- aov(PC2 ~ Genotype, data=pcadata)
summary(aov2) 

aov3 <- aov(PC3 ~ Genotype, data=pcadata)
summary(aov3) 

aov4 <- aov(PC4 ~ Genotype, data=pcadata)
summary(aov4) 


pcadata$Genotype <- factor(pcadata$Genotype, levels=c("WT", "FMR1"))

PCA12 <- plotPCs(pcadata, 1, 2, aescolor = pcadata$Genotype, colorname = " ", aesshape = pcadata$Genotype, shapename = " ",  colorvalues = colorvalGenotype)
PCA12

PCA34 <- plotPCs(pcadata, 3, 4,  aescolor = pcadata$Genotype, colorname = " ", aesshape = pcadata$Genotype, shapename = " ",  colorvalues = colorvalGenotype)
PCA34

# pdf the same pca plots descripbed above of the above
pdf(file="../figures/02_RNAseq/PCA12.pdf", width=3, height=2.5)
plot(PCA12)
dev.off()

pdf(file="../figures/02_RNAseq/PCA34.pdf", width=3, height=2.5)
plot(PCA34)
dev.off()

```

# Number of differentially expressed genes per two-way contrast

```{r Twowaycontrasts3}
#calculate significance of all two way comparisions
# see source "functions_RNAseq.R" 

contrast1 <- resvals(contrastvector = c("Genotype", "FMR1", "WT"), mypval = 0.05) # 8
contrast2 <- resvals(contrastvector = c("Genotype", "FMR1", "WT"), mypval = 0.2) # 20
contrast3 <- resvals(contrastvector = c("Genotype", "FMR1", "WT"), mypval = 0.1) # 11


# gene list
res <- results(dds, contrast =c("Genotype", "FMR1", "WT"), independentFiltering = F, alpha = 0.2)
resOrdered <- res[order(res$padj),]
head(resOrdered, 20)
data <- data.frame(gene = row.names(res), padj = (res$padj), lfc = res$log2FoldChange)
data <- na.omit(data)
data <- filter(data, padj < 0.1)
data[order(data$padj),]




res <- results(dds, contrast =c("Genotype", "FMR1", "WT"), independentFiltering = T, alpha = 0.05)
summary(res)

topGene <- rownames(res)[which.min(res$padj)]
plotCounts(dds, gene = topGene, intgroup=c("Genotype"))

resOrdered <- res[order(res$padj),]
head(resOrdered, 10)

data <- data.frame(gene = row.names(res),
                   pvalue = -log10(res$padj), 
                   lfc = res$log2FoldChange)
data <- na.omit(data)
data <- data %>%
  mutate(color = ifelse(data$lfc > 0 & data$pvalue > 1.3, 
                        yes = "FMR1", 
                        no = ifelse(data$lfc < 0 & data$pvalue > 1.3, 
                                    yes = "WT", 
                                    no = "none")))
top_labelled <- top_n(data, n = 5, wt = lfc)


# Color corresponds to fold change directionality
colored <- ggplot(data, aes(x = lfc, y = pvalue)) + 
  geom_point(aes(color = factor(color)), size = 0.8, alpha = 0.5, na.rm = T) + # add gene points
  theme_bw(base_size = 8) + # clean up theme
  theme(legend.position = "none") + # remove legend 
  scale_color_manual(values = volcanoGenotype) + 
  theme(panel.grid.minor=element_blank(),
        panel.grid.major=element_blank()) + 
  scale_x_continuous(name="log2 (consistent/yoked)",
                     limits=c(-2, 2)) +
  scale_y_log10(name="-log10 (adjusted p-value)") +
  geom_hline(yintercept = 1.3,  size = 0.25, linetype = 2 ) +
  #geom_text(data=filter(data, pvalue>1.4), aes(label=gene), size = 1, nudge_x = 0.4) +
  geom_text(data=filter(data, pvalue>1.4), aes(label=gene), size = 0.8, nudge_x = 0.3)
colored

pdf(file="../figures/02_RNAseq/volcano.pdf", width=1.5, height=1.75)
plot(colored)
dev.off()


```

```{r heatmap, echo=F}
DEGes <- assay(rld)
DEGes <- cbind(DEGes, contrast2)
DEGes <- as.data.frame(DEGes) # convert matrix to dataframe
DEGes$rownames <- rownames(DEGes)  # add the rownames to the dataframe
DEGes$padjmin <- DEGes$padjGenotypeFMR1WT

# create new col with min padj
DEGes <- DEGes %>% filter(padjmin < 0.2)
rownames(DEGes) <- DEGes$rownames
drop.cols <-colnames(DEGes[,grep("padj|pval|rownames", colnames(DEGes))])
DEGes <- DEGes %>% dplyr::select(-one_of(drop.cols))
DEGes <- as.matrix(DEGes)
DEGes <- DEGes - rowMeans(DEGes)

## the heatmap annotation file
df <- as.data.frame(colData(dds)[,c("Genotype")]) ## matrix to df
rownames(df) <- names(countData)

ann_colors <- ann_colorsGenotype # see color options 

# make sure the data is a matrix
DEGes <- as.matrix(DEGes) 

# set color breaks
paletteLength <- 30
myBreaks <- c(seq(min(DEGes), 0, length.out=ceiling(paletteLength/2) + 1), 
              seq(max(DEGes)/paletteLength, max(DEGes), length.out=floor(paletteLength/2)))

pheatmap(DEGes, show_colnames=F, show_rownames = T,
         annotation_col=df, annotation_colors = ann_colors,
         treeheight_row = 0, treeheight_col = 25,
         fontsize = 11, 
         #width=4.5, height=3,
         border_color = "grey60" ,
         color = viridis(30),
         cellwidth = 8, 
         clustering_method="average",
         breaks=myBreaks,
         clustering_distance_cols="correlation" 
         )

# for adobe
pheatmap(DEGes, show_colnames=F, show_rownames = F,
         annotation_col=df, annotation_colors = ann_colors,
         treeheight_row = 0, treeheight_col = 50,
         fontsize = 10, 
         #width=4.5, height=3,
         border_color = "grey60" ,
         color = viridis(30),
         cellwidth = 8, 
         clustering_method="average",
         breaks=myBreaks,
         clustering_distance_cols="correlation",
         filename = "../figures/02_RNAseq/pheatmap.pdf"
         )

# no legends just the heatmap
pheatmap(DEGes, show_colnames=F, show_rownames = T,
         annotation_col=df, annotation_colors = ann_colors, 
         annotation_row = NA, 
         annotation_legend = FALSE,
         annotation_names_row = FALSE, annotation_names_col = FALSE,
         treeheight_row = 0, treeheight_col = 50,
         fontsize = 11, 
         border_color = "grey60" ,
         color = viridis(30),
         #cellwidth = 7, 
         clustering_method="average",
         breaks=myBreaks,
         clustering_distance_cols="correlation" 
         )

pheatmap(DEGes, show_colnames=F, show_rownames = T,
         annotation_col=df, annotation_colors = ann_colors, 
         annotation_row = NA, 
         annotation_legend = FALSE,
         annotation_names_row = FALSE, annotation_names_col = FALSE,
         treeheight_row = 0, treeheight_col = 50,
         fontsize = 11, 
         border_color = "grey60" ,
         color = viridis(30),
         #cellwidth = 7, 
         clustering_method="average",
         breaks=myBreaks,
         clustering_distance_cols="correlation", 
         filename = "../figures/02_RNAseq/pheatmap_minimal.pdf"
         )
```


```{r heatmap05, echo=F}
DEGes <- assay(rld)
DEGes <- cbind(DEGes, contrast1)
DEGes <- as.data.frame(DEGes) # convert matrix to dataframe
DEGes$rownames <- rownames(DEGes)  # add the rownames to the dataframe
DEGes$padjmin <- DEGes$padjGenotypeFMR1WT

# create new col with min padj
DEGes <- DEGes %>% filter(padjmin < 0.05)
rownames(DEGes) <- DEGes$rownames
drop.cols <-colnames(DEGes[,grep("padj|pval|rownames", colnames(DEGes))])
DEGes <- DEGes %>% dplyr::select(-one_of(drop.cols))
DEGes <- as.matrix(DEGes)
DEGes <- DEGes - rowMeans(DEGes)

## the heatmap annotation file
df <- as.data.frame(colData(dds)[,c("Genotype")]) ## matrix to df
rownames(df) <- names(countData)

ann_colors <- ann_colorsGenotype # see color options 

# make sure the data is a matrix
DEGes <- as.matrix(DEGes) 

# set color breaks
paletteLength <- 30
myBreaks <- c(seq(min(DEGes), 0, length.out=ceiling(paletteLength/2) + 1), 
              seq(max(DEGes)/paletteLength, max(DEGes), length.out=floor(paletteLength/2)))

pheatmap(DEGes, show_colnames=F, show_rownames = T,
         annotation_col=df, annotation_colors = ann_colors,
         treeheight_row = 0, treeheight_col = 25,
         fontsize = 11, 
         #width=4.5, height=3,
         border_color = "grey60" ,
         color = viridis(30),
         cellwidth = 8, 
         clustering_method="average",
         breaks=myBreaks,
         clustering_distance_cols="correlation" 
         )


pheatmap(DEGes, show_colnames=F, show_rownames = T,
         annotation_col=df, annotation_colors = ann_colors, 
         annotation_row = NA, 
         annotation_legend = FALSE,
         annotation_names_row = FALSE, annotation_names_col = FALSE,
         treeheight_row = 0, treeheight_col = 50,
         fontsize = 11, 
         border_color = "grey60" ,
         color = viridis(30),
         #cellwidth = 7, 
         clustering_method="average",
         breaks=myBreaks,
         clustering_distance_cols="correlation", 
         filename = "../figures/02_RNAseq/pheatmap_minimal05.pdf"
         )
```



## Write the two files

```{r writecsv}
write.csv(vsd, file = "../data/02_vsd.csv", row.names = T)
write.csv(rlddf, file = "../data/02_rlddf.csv", row.names = T)
write.csv(colData, file = "../data/02_colData.csv", row.names = T)
```